MS_data_clean

### 2022-07-20 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
genus level push

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% select(-X)

@@

@@

```{r, fig.width=5,fig.height=2}
beeCiteNoFilter <- interactions_b %>% select(sourceTaxonSpeciesName, targetTaxonFamilyName) %>% dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% group_by(interestBee, interact) %>% count() %>% ungroup() %>% filter(interestBee != "", interact != "")

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(n)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(n))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 5 citations to contribute to degree

beeIntDegreeW <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeW$numcitations <- beeCiteNoFilter$n
beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## weighted degree

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type = 'inner')

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee",  type = 'inner')

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee",  type = 'inner')

beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee",  type = 'inner')

beeCiteW
```

@@

western_table$west_specialist <- 1
western_test <- western_table %>% dplyr::select(interestBee, west_specialist) %>% unique()

eastern_table$east_specialist <- 1
eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist) %>% unique()

central_table$cent_specialist <- 1
central_test <- central_table %>% dplyr::select(interestBee, cent_specialist) %>% unique()

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0

@@

fit = glm(specialist ~ n_Bee + degree, data = plotdf, family = 'binomial')

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 40))+ geom_jitter()+ scale_color_discrete(name = "Specialist") + geom_smooth()

@@

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist)), text = paste("Bee:", interestBee)) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity (Minimum 5 Citations) and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 15))+ geom_jitter() + scale_color_discrete(name = "Specialist")+ geom_smooth(se = FALSE)

@@

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Function of Degree')) + ggtitle(paste('Function of Degree and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 1)) + geom_smooth(se = FALSE)+ scale_color_discrete(name = "Specialist")

@@

>>>>

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)

```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% dplyr::select(-X)

@@

```{r}
interactions_b[1:5,]
```

@@

```{r, fig.width=5,fig.height=2}
beeCiteNoFilter <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName,targetTaxonFamilyName) %>% 
  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% dplyr::group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
  filter(interestBee != "", interact != "")

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup() %>% unique()

beeIntDegree <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count() 
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()

beeIntDegreeC10 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree

beeIntDegreeW <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeW$numcitations <- beeCiteNoFilter$n
beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree

beeCite <-  left_join(beeCiteNoFilter, beeIntDegree, by = 'interestBee')

beeCiteC10 <-  left_join(beeCiteNoFilter, beeIntDegreeC10, by = 'interestBee')

beeCiteC5 <-  left_join(beeCiteNoFilter, beeIntDegreeC5, by = 'interestBee')

beeCiteW <-  left_join(beeCiteNoFilter, beeIntDegreeW, by = 'interestBee')

beeCiteNoFilterGenus <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName, targetTaxonGenusName) %>% 
  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonGenusName) %>% group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
  filter(interestBee != "", interact != "")

TotalCitationsGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
TotalCitationsGenus <- sum(TotalCitationsGenus$n_Bee)

beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegreeGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeGenus <- beeIntDegreeGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()

beeIntDegreeC10Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeC10Genus$numcitations <- beeCiteNoFilterGenus$n
beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% filter(numcitations >= 10)
beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree

beeIntDegreeC5Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeC5Genus$numcitations <- beeCiteNoFilterGenus$n
beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% filter(numcitations >= 5)
beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree

beeIntDegreeWGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
beeIntDegreeWGenus$numcitations <- beeCiteNoFilterGenus$n
beeIntDegreeWGenus$n_Bee <- beeCiteNoFilterGenus$n_Bee
beeIntDegreeWGenus$n <- (beeIntDegreeWGenus$numcitations / beeIntDegreeWGenus$n_Bee)^2
beeIntDegreeWGenus <- beeIntDegreeWGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree

beeCiteGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeGenus, by = 'interestBee')

beeCiteC10Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC10Genus, by = 'interestBee')

beeCiteC5Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC5Genus, by = 'interestBee')

beeCiteWGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeWGenus, by = 'interestBee')

beeCite[1:10,]
```

```{r}
beeCiteGenus <- beeCiteGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeGenus = degree)

beeCite <- beeCite %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeFamily = degree)

beeCiteC5Genus <- beeCiteC5Genus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Genus = degree)

beeCiteC5 <- beeCiteC5 %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Family = degree)

beeCiteWGenus <- beeCiteWGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWGenus = degree)
beeCiteW <- beeCiteW %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWFamily = degree)

beeCiteW[1:5,]
```

@@

western_table$west_specialist <- 1
western_test <- western_table %>% dplyr::select(interestBee, west_specialist, Status) %>% dplyr::rename('Status_W' = 'Status') %>% unique()

eastern_table$east_specialist <- 1
eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist, Status)%>% dplyr::rename('Status_E'= 'Status') %>% unique()

central_table$cent_specialist <- 1
central_test <- central_table %>% dplyr::select(interestBee, cent_specialist, Status)%>% dplyr::rename('Status_C'= 'Status') %>% unique()

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)

@@

@@

@@

@@

```{r}
plotdf <- merge(beeCiteGenus, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

write.csv(plotdf, 'BeeCiteRawGenus')
```

```{r}
plotdf <- merge(beeCiteC5Genus, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

write.csv(plotdf, 'BeeCiteC5Genus')
```

```{r}
plotdf <- merge(beeCiteWGenus, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

write.csv(plotdf, 'BeeCiteWGenus')
```

### 2022-07-20 MS_data_modeling/MS_data_explore.Rmd
genus level push

globee_T_formatGenus <- globee_TGenus %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_TGenus = degreeWGenus) %>% 
  dplyr::select(interestBee, degreeGLOBEE_TGenus) %>% unique() %>%
  arrange(interestBee)%>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)  
compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), all = T) 
compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T) 
compare_df <- merge(compare_df, globee_formatGenus, by = c('interestBee', 'number_of_citation_for_bee_species'), all = T)  
compare_df <- merge(compare_df, globee_5cut_formatGenus, by = c('interestBee'), all = T) 
compare_df <- merge(compare_df, globee_T_formatGenus, by = c('interestBee'), all = T)

@@

final <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% 
  dplyr::rename(interest_bee_species = interestBee, number_of_citations_for_bee_species_GloBi = number_of_citation_for_bee_species, degree_raw_GloBi_family = degreeGLOBEEFamily,
                pollen_hosts_GloBi_family = pollenHostGLOBEEFamily, pollen_host_genus_MS = pollenHostsMS, degree_MS_family = degreeMS, degree_MS_Genus = degreeMSGenus,
                specialization_MS = specializationMS, References_MS = ReferencesMS, pollen_host_family_MS = pollenHostsFamilysMS, 
                degree_cut_5_GloBi_family = degreeGLOBEE_C5Family, degree_transformed_GloBi_family = degreeGLOBEE_TFamily, 
                degree_raw_GloBi_genus = degreeGLOBEEGenus, pollen_hosts_GloBi_genus = pollenHostGLOBEEGenus, degree_cut_5_GloBi_genus = degreeGLOBEE_C5Genus,
                degree_transformed_GloBi_genus = degreeGLOBEE_TGenus, specialist_FOWLER = specialistFOWLER) %>% 
  dplyr::select(interest_bee_species, number_of_citations_for_bee_species_GloBi, degree_raw_GloBi_family, degree_cut_5_GloBi_family, degree_transformed_GloBi_family,
                pollen_hosts_GloBi_family, degree_MS_family, pollen_host_family_MS, degree_raw_GloBi_genus, degree_cut_5_GloBi_genus, degree_transformed_GloBi_genus,
                pollen_hosts_GloBi_genus, pollen_host_genus_MS, specialization_MS) %>% 
           mutate(pollen_host_family_MS = str_remove(pollen_hosts_GloBi_family, "NA, "),
                  pollen_hosts_GloBi_family = str_to_title(pollen_hosts_GloBi_family), pollen_host_family_MS =str_to_title(pollen_host_family_MS),
                  pollen_hosts_GloBi_genus = str_to_title(pollen_hosts_GloBi_genus))

@@

write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/specialization_degree_df')
```

>>>>

globee_T_formatGenus <- globee_TGenus %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_TGenus = degreeWGenus) %>% 
  dplyr::select(interestBee, degreeGLOBEE_TGenus) %>% unique() %>%
  arrange(interestBee)%>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))

globee_formatGenus %>% filter(interestBee == 'Andrena Accepta')
globee_5cut_formatGenus %>% filter(interestBee == 'Andrena Accepta')
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), type = 'all')  
compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), type = 'all') 
compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), type = 'all') 
compare_df <- merge(compare_df, globee_formatGenus, by = c('interestBee', 'number_of_citation_for_bee_species'), type = 'all') 
compare_df <- merge(compare_df, globee_5cut_formatGenus, by = c('interestBee'), type = 'all') 
compare_df <- merge(compare_df, globee_T_formatGenus, by = c('interestBee'), type = 'all')

@@

compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% dplyr::select(interestBee, degreeGLOBEEGenus, degreeGLOBEE_C5Genus) %>%
  filter(!is.na(degreeGLOBEE_C5Genus), is.na(degreeGLOBEEGenus))

final <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% 
  dplyr::rename(interest_bee_species = interestBee, number_of_citations_for_bee_species_GloBi = number_of_citation_for_bee_species, degree_raw_GloBi_family = degreeGLOBEEFamily,
                visitation_hosts_GloBi_family = pollenHostGLOBEEFamily, pollen_host_genus_MS = pollenHostsMS, degree_MS_family = degreeMS, degree_MS_Genus = degreeMSGenus,
                specialization_MS = specializationMS, References_MS = ReferencesMS, pollen_host_family_MS = pollenHostsFamilysMS, 
                degree_cut_5_GloBi_family = degreeGLOBEE_C5Family, degree_transformed_GloBi_family = degreeGLOBEE_TFamily, 
                degree_raw_GloBi_genus = degreeGLOBEEGenus, visitation_hosts_GloBi_genus = pollenHostGLOBEEGenus, degree_cut_5_GloBi_genus = degreeGLOBEE_C5Genus,
                degree_transformed_GloBi_genus = degreeGLOBEE_TGenus, rarity_east_fowler = Status_E, rarity_central_fowler = Status_C, rarity_west_fowler = Status_W,
                degree_MS_genus = degreeMSGenus) %>% 
  dplyr::select(interest_bee_species, number_of_citations_for_bee_species_GloBi, degree_raw_GloBi_family, degree_cut_5_GloBi_family, degree_transformed_GloBi_family,
                visitation_hosts_GloBi_family, degree_MS_family, pollen_host_family_MS, degree_raw_GloBi_genus, degree_cut_5_GloBi_genus, 
                degree_transformed_GloBi_genus, degree_MS_genus,
                visitation_hosts_GloBi_genus, pollen_host_genus_MS, specialization_MS, rarity_east_fowler, rarity_central_fowler, rarity_west_fowler) %>% 
           mutate(pollen_host_family_MS = str_remove(visitation_hosts_GloBi_family, "NA, "),
                  visitation_hosts_GloBi_family = str_to_title(visitation_hosts_GloBi_family), pollen_host_family_MS =str_to_title(pollen_host_family_MS),
                  visitation_hosts_GloBi_genus = str_to_title(visitation_hosts_GloBi_genus))

@@

write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/visitation_specialization_df')
```
